---
title: "Intro"
subtitle: "Week One"
format:
  pptx:
    incremental: true   
    reference-doc: templates/SOC2069_pptx_template.pptx
---

# In the morning

## Getting up

- Turn off alarm
- Get out of bed
  - something else
  - something other

## Breakfast

- Eat eggs
- Drink coffee
- more eggs
- loads of coffee

# In the evening



## Dinner

- Eat spaghetti
- Drink wine

## Going to sleep

- Get in bed
- Count sheep

## Incremental stuff

:::: {.columns}

::: {.column width="40%"}
contents...
:::

::: {.column width="60%"}
contents...
:::

::::

::: {.notes}
I may write a lot of extra stuff here in case you are interested.
:::


## Title

```{r}

library(tidyverse)
library(palmerpenguins)
```


```{r}
#| echo: true

ggplot(penguins,                        
       aes(x = bill_length_mm,          
           y = bill_depth_mm, 
           color = species)) +
  geom_point()
```

## 

```{r}
#| echo: true
#| output-location: column-fragment
#| code-line-numbers: "|2"

library(ggplot2)

mtcars |> 
  ggplot(aes(x = disp, y = mpg)) +
  geom_point() +
  geom_smooth(method = "loess", formula = "y~x")
```


## Final

This is an [important sentence!]{.fragment .highlight-red}

