Working through this example: https://library.virginia.edu/data/articles/a-beginners-guide-to-marginal-effects

# A Beginner's Guide to Marginal Effects

## Data

```{r}
set.seed(1)
x <- sort(runif(20, -5, 10))
y <- 1.5 + 3*x - 0.5*x^2 + rnorm(20, sd = 3)
d <- data.frame(y = y, x = x)
plot(y ~ x, data = d)
```

## Model

```{r}
m <- lm(y ~ x + I(x^2), data = d)
summary(m)
```

## Plot fitted line

```{r}
plot(y ~ x, data = d)
lines(x, fitted(m))
```

## Average marginal effects" with `{margins}`

```{r}
library(margins)
mfx <- margins(m, variables = "x")
summary(mfx)
```

## "Average marginal effects" using other functions

## Derivatives

The derivative of a function is a fundamental concept in calculus that represents how a function changes as its input changes.

The definition of the derivative is:

$$ f'(x) = \lim_{{h \to 0}} \frac{{f(x + h) - f(x)}}{h} $$

Let's unpack the definition of a derivative step by step:

1.  Average Rate of Change:

For a function $f(x)$, the average rate of change between two points $(x, f(x))$ and $(x + h, f(x + h))$ is given by the formula: $$ \text{Average Rate of Change} = \frac{f(x + h) - f(x)}{h} $$

Here, $h$ represents the change in the input variable $x$.
It's often used in the definition of derivatives to denote an infinitesimally small change in the input variable (in this case, "x").

When calculating derivatives, you're essentially finding the rate at which a function changes as the input (usually denoted by "x") changes by a very small amount "h." Taking the limit as "h" approaches 0 gives you the instantaneous rate of change at a specific point.

The symbol "h" is used traditionally, but you might also see other letters like $\delta x$ or $\Delta x$ (i.e. lower-case or capital Greek letter *delta*; *delta* is the initial letter of the Greek word διαφορά \[diaphorá\], meaning "difference") used in similar contexts, where they all serve a similar purpose of denoting a small change in the input variable.

2.  Instantaneous Rate of Change:

The derivative represents the instantaneous rate of change of a function at a specific point.
To find this instantaneous rate of change, we need to let the interval $\Delta x$ (or $h$) approach 0.
This is [expressed mathematically as a limit]{.underline}: $$ f'(x) = \lim_{{h \to 0}} \frac{f(x + h) - f(x)}{h} $$

3.  Interpretation:

-   **Geometrically:** The derivative at a point $x$ represents the slope of the tangent line to the graph of the function at that point. It indicates how steeply the function is rising or falling at that specific point.
-   **Physically:** In physics, the derivative often represents velocity when the function represents displacement with respect to time. For example, if $s(t)$ represents the position of an object at time $t$, then $s'(t)$ represents the velocity of the object at time $t$.

4.  Differentiability:

A function is said to be differentiable at a point if the derivative exists at that point.
Not all functions are differentiable everywhere; there are certain conditions and types of functions that are differentiable only within specific intervals or points.

5.  Notation:

The derivative of a function $f(x)$ is denoted as $f'(x)$ or $\frac{df}{dx}$.
Higher order derivatives, indicating how the rate of change itself changes, can be denoted as $f''(x)$ (second derivative), $f'''(x)$ (third derivative), and so on.

## The derivative of `y` with respect to `x`: the *power rule*

We have the following function resulting from a fitted regression model:

$$
y = 2.25 + 2.98x - 0.51x^2
$$

To find its derivative with respect to $x$, we differentiate each term using the power rule.

The power rule states that the derivative of a function of the form $f(x) = x^n$, where $n$ is a constant, with respect to $x$ is given by:

$$
\frac{d}{dx} (x^n) = n \times x^{(n-1)}
$$

In words, to differentiate a term where $x$ is raised to a constant power $n$, you bring down the exponent as the coefficient and then decrease the exponent by 1.

To differentiate the equation above:

1.  The derivative of the constant term $2.25$ is $0$ because the derivative of a constant is always $0$.
    This results from the definition of a "derivative" (see above).
    For a constant function, $f(x) = c$, where $c$ is a constant, $f(x + h) = c$ for any $h$.
    Therefore, the expression becomes: $\frac{{c - c}}{h} = \frac{0}{h} = 0$.
    As $h$ approaches 0, the fraction $\frac{0}{h}$ remains 0.
    Therefore, the limit as $h$ approaches 0 is 0.
    This proves that the derivative of the constant function $f(x) = c$ is always 0 for any constant $c$.

2.  For the term $2.98x$, the derivative with respect to $x$ is $2.98$ because it follows the form $ax^1$ (where $a = 2.98$).

3.  For the term $-0.51x^2$, where $-0.51$ is a constant and $x^2$ follows the form $x^n$ with $n = 2$, we therefore get:

$$
\begin{split}
\frac{d}{dx}(-0.51x^2) &= 2 \times (-0.51) \times x^{2-1} \\ &= 2 \times (-0.51) \times x^1 \\ &=  2 \times (-0.51) \times x \\ &= -1.02 \times x
\end{split}
$$

So, the derivative of ($-0.51x^2$) with respect to $x$ is $-1.02x^{(2-1)} = -1.02x$ using the power rule.

Summing up these derivatives, the derivative of the function $y = 2.25 + 2.98x - 0.51x^2$ with respect to $x$ is:

$$
\begin{split}
\frac{dy}{dx} &= 0 + 2.98 + 2(-0.51)x \\ &= 0 + 2.98 - 1.02x
\end{split}
$$

So to calculate the "dydx" values we can do the following (showing only the first six values):

```{r}
# m$coefficients[2] = 2.98
# m$coefficients[3] = -0.51
head(m$coefficients[2] + 2 * m$coefficients[3] * d$x)
```

We can confirm these match the "dydx" column in the mfx data frame:

```{r}
head(mfx$dydx_x)
```

Now let's take the average of these "dydx" values.

```{r}
mean(mfx$dydx_x)
```

And there it is, our average marginal effect.
It is literally the average of the marginal effects.
Stated another way, it is the average of the slopes of the tangent lines at 20 different xy coordinates.
Is this useful?
Probably not in this case.
The effect of x changes as itself changes.
The average of all these effects doesn't tell us much.
We're better off just looking at our original plot with the fitted line.

A better approach may be to examine marginal effects at representative values.
For example, what if we were interested in the marginal effects at x = -1 and x = 6?
We can use the at argument to specify at which x values to calculate the marginal effects.
These need to specified as a list object.
If you're a Stata user, the equivalent code would be margins, dydx(x) at(x = (-1 6)).

```{r}
mfx2 <- margins(m, variables = "x", 
                at = list(x = c(-1, 6)))
summary(mfx2)
```

Now we have two measures of the effect of "x".
When x = -1, we can imagine a tangent line with a positive slope of about 4.
When x = 6, we can imagine a tangent line with a negative slope of about -3.2.
The summary output provides 95% confidence intervals on these estimates in the "lower" and "upper" columns.

# Now let's look at marginal effects from another angle, this time in the context of more realistic data.
