---
title: "Timetable"
format: 
  html: 
    grid:
      sidebar-width: 100px
      margin-width: 100px
      body-width: 1100px
---




```{r, echo=F}


# Set dates
cal_start <- as.Date("2023/09/1")
mod_start <- as.Date("2023/09/25")
mod_break_1_start <- as.Date("2023/11/06")
mod_break_1_end <- as.Date("2023/11/20")
mod_break_2_start <- as.Date("2023/12/18")
mod_break_2_end <- as.Date("2024/01/01")
mod_end <- as.Date("2024/01/14")
cal_end <- as.Date("2024/01/30")

firstHalf <- seq(mod_start, mod_break_1_start, by="1 day")
secondHalf <- seq(mod_break_1_end, mod_break_2_start, by="1 day")
wholeMod <- seq(mod_start, mod_end, by="1 day")
wholeCal <- seq(cal_start, cal_end, by="1 day")


# View(data.frame(wholeMod))
# View(data.frame(wholeCal))


assessment_1 <- match(as.Date("2023/11/24"), wholeCal)
assessment_2 <- match(as.Date("2024/01/12"), wholeCal)

lectures <- c(seq(mod_start+1, mod_break_1_start, by="7 day"),
              seq(mod_break_1_end+1, mod_break_2_start, by="7 day"))

lectures_num <- match(lectures, wholeCal)



highlights <- rep(NA, length(wholeCal))

highlights[lectures_num] <- "Lecture"
highlights[assessment_1] <- "Assignment 1"
highlights[assessment_2] <- "Assignment 2"


calendR::calendR(start_date = "2023/09/1", end_date = "2024/01/30", 
                 title = "SOC2069 23/24",
                 orientation = "portrait", 
                 start = "M",
                 mbg.col = 4,                                                    # Background colour for the month names
                 months.col = "white",                                           # Text colour of the month names
                 special.days = highlights,
                 special.col = c("pink", "pink", "lightgreen"),                   # Colour of the special.days
                 months.pos = 0.5,                                               # Horizontal alignment of the month names)
                 legend.pos = "right"
                 ) 
```


```{r}
#| eval: false
#| echo: false

t <- calendR::calendR(year = 2021, month = 8, papersize = "A0")

```


There are also six **IT lab** groups taking place at various times and locations between Tuesdays and Fridays. Make sure to check your timetable app to see which group you have been assigned to, and when and where your IT lab will take place each week. A full module timetable [is available here](https://timetables.ncl.ac.uk/view.php?timetable=modules&show=grid&period=sem1&type=C%3BF%3BH%3BL%3BP%3BS%3BT%3BV%3BW%3BA%3BSO&modules=SOC2069&programme=&stage=&studentId=). The labs last 1.5 hours and are run by two teaching assistants. In the labs, you will be working through the exercises on the worksheets.



