---
title: ""
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
self-contained: true
---

```{r message=FALSE, warning=FALSE, include=FALSE}
## To make your lives easier, I include this code chunk that installs and loads all the packages that we have used this semester, and which you may want to use for the assignment

## Because we do not want any information about this to actually get printed into the output document, I have added some options to the code-chunk header above

## For this reason, leave this chunk as it is. All the other code chunks that you will include later should print everything in the output (code and results)

## To make sure that the packages listed below are loaded into your session (so that you can run your code chunks below separately to keep track of your results), run this code chunk first by clicking on the green righ-pointing arrow at the top-right corner of this code chunk

options(repos = list(CRAN = "http://cran.rstudio.com/")) # Just tells R where to download the packages from

install.packages("tidyverse")
install.packages("mosaic")
install.packages("sjmisc")
install.packages("jtools")

library(tidyverse)
library(mosaic)
library(sjmisc)
library(jtools)
```



# My title (as a first-level heading)

All of the text in this document can **and should** be deleted. They are just there to give you some guidance. You will write your own text here and can choose to keep or replace the headings (text that begins with # hashtags) too.


## Introduction (as a second-level heading)

This is where I introduce the topic of my assignment, stating what I aim to find out and what my research question is.

## Literature review

This is where I place my research question in the context of existing literature, summarizing what we already know about the topic based on a few academic sources (around 5 should be enough for this assignment).


## Data and methods

This is where I describe the data source that I am using to answer the research question. The data will be the same Wave 8 of the UK Household Longitudinal Study (Understanding Society) survey dataset that we have been using in the labs. Here, I'll say a few words about that data based on the information available on the survey website, and I will provide a citation of the dataset (it's all on Canvas).

I also introduce the variables that I have selected to answer the research question. I briefly describe what they measure and whether I will transform them in any way before using them in the analysis.

I also briefly state the methods I will use (what type of regression analysis, any tables/figures I include, etc.)

## Results

This is where I present my analysis, step-by-step, as I have done in the labs. This section will include:
- the code I use to produce the outputs
- the outputs that are produced
- my brief interpretation of the meaning of the outputs

To include a code chunk that will be executed as R code, I either type **ctrl** + **alt** + **i** (or **Command** + **Option** + **i** on macOS), or I click on the ***+C*** icon on this document's menu bar and select "R". The code chunk will look like this:

```{r}
# I can now use this space just like an R script editor, as I've done in the labs; this line, for example, is just a comment because it begins with a #

# For example, as a first step, load the dataset:


ukhls <- "blablabla... replace this, obviously"


# Don't forget to delete these comments before you "knit" the document to Microsoft Word or html; you can add your own comments if you want to include some information about what your code is aimed to do

```

It is advisable to include several code chunks, one for each command of the analysis, interspersed with your interpretation of the results.

While you work on the document, you can execute each command separately, one by one, to get and check the results. 

You may also want to work in an `R` script first if you prefer, to get the results, then cut and paste the code into chunks here (but this is additional and inefficient work, so I do not recommend it!)



## Discussion and conclusions

This is where I take the broader picture of what my analysis and outputs have shown and connect it back to the existing literature, stating how my findings relate to what we already know.





## References

Here I include a reference list for the literature that I cited above.

Once everything is finished and I'm satisfied with the content, I can "knit" this document to Microsoft Word (or html) to get everything exported to a usable format.



