---
title: ""
format: 
  docx: default
    df_print: kable
execute: 
  warning: false
  error: false
  echo: true
---

```{r include=FALSE}

# Load packages
# The `include=FALSE` setting above ensures that no output from this code chunk will be shown in the rendered Word document

if (!require("pacman")) install.packages("pacman")    # install {pacman} if not yet installed

pacman::p_load(tidyverse,             # metapackage; general data management (especially the `dplyr` package)
               easystats,             # metapackage; data management and model summaries
               gtsummary,             # data and model summary tables
               ggformula,             # formula syntax to simplify `ggplot2` graphics
               sjlabelled,            # work with labelled survey data from other statistical packages
               ggeffects,             # extract and plot marginal/conditional/interaction effects from models
               marginaleffects,       # extract and plot marginal/conditional/interaction effects from models
               flextable              # render tables nicely in MS Word/other text editors 
               )
```

# My main title (as a first-level heading)

All the text in this document should be deleted.
It is here only to give you some guidance.
You will write your own text and headings.

## Introduction (as a second-level heading)

This is where I introduce the topic of my assignment, stating what I aim to find out and what my research question is.

## Literature review

This is where I place my research question in the context of existing literature, summarizing what we already know about the topic based on a few academic sources (around 5 references related to the substantive topic of the research question should be enough for this assignment; references to the dataset or methodology should be additional).

## Data and methods

This is where I describe the data source that I am using to answer the research question.
Make sure to consult the survey documentation in order to give a very brief description of the survey itself, then describe the variables you have chosen for analysis.
The dataset should be cited as advised on the data description webpage.

I also introduce the variables that I have selected to answer the research question.
I briefly describe what they measure and whether I have transformed them in any way before using them in the analysis.

I also briefly state the methods I will use.

I can also include any descriptive tables/graphs in this section if I want.
To include a code chunk that will be executed as R code, I can either type **ctrl** + **alt** + **i** (or **Command** + **Option** + **i** on macOS), or I click on the ***+C*** icon on this document's menu bar and select "R".
The code chunk will look like this:

```{r}

```

There are several useful code chunk settings that I can use.
By default, the code and all output from a code-chunk will be rendered to the Word document.
However, I may want to have a code-chunk that doesn't render any output, for example if I want to produce some tables only for checking variables, but not worthy of including in the final report.
For that purpose, I can use the `include = FALSE` option in the code chunk's header, like below:

```{r include=FALSE}

# Example: nothing from this chunk will be printer to the Word document

head(cars)

```

I can also set a chunk so that the code gets rendered to Word, but the output from the code does not, using the `eval = FALSE` setting.
This can be useful if you want to demonstrate your coding, but the output itself is not meant to be included in the final report:

```{r eval=FALSE}

# Example: only the code will be printed to the Word document

head(cars)

```

## Results

This is where I present my analysis.
This section will include:

-   the code I use to produce the outputs

-   the outputs that are produced

-   my brief interpretation of the meaning of the outputs

```{r}



```

It is advisable to include several code chunks, one for each command of the analysis, interspersed with your interpretation of the results.

While you work on the document, you can execute each command separately, one by one, to get and check the results.
I can then decide whether you do not want to include the outputs from certain chunks, and add the required chunk header settings, as shown above.

## Discussion and conclusions

This is where I take the broader picture of what my analysis and outputs have shown and connect it back to the existing literature, stating how my findings relate to what we already know.

## References

Here I include a reference list for the literature that I cited above.

Once everything is finished and I'm satisfied with the content, I can "Render" this document to Microsoft Word.
I can then further edit the document directly in Word.
