---
title: "Introduction: Information, Data, Models"
subtitle: "Week 1"
format:
  pptx:
    incremental: true   
    reference-doc: templates/SOC2069_pptx_template.pptx
---

```{r}

# Packages
library(tidyverse)
library(palmerpenguins)
```


# Information > Data > Models



## Models in social science

- Descriptive models
- Predictive models
- Inferential models
- Causal models




## Typical research objectives by required sample size (*n*)

- **$n\to\infty$**
- **n $\approx$ 5,000 - 100,000**: stratified population survey
- **n $\approx$ 1,000 - 5,000**: population survey
- **n $\approx$ 100 - 1,000**: targeted survey
- **n $\approx$ 50 - 100**: stratified narrative study
- **n $\approx$ 10 - 50**: narrative study
- **n $\approx$ 5 - 10**: interactive narrative study
- **n = 2 - 4**: comparative case study
- **n = 1**: case study
- **n = 0**: theoretical/simulation study



--------------------------------------------------------------------------------


# Module outline



# Application: social trust



## Dinner

- Eat spaghetti
- Drink wine

## Going to sleep

- Get in bed
- Count sheep

## This print "Two content" output

:::: {.columns}

::: {.column width="40%"}
```{r}
#| echo: true
#| eval: true

ggplot(penguins,                        
       aes(x = bill_length_mm,          
           y = bill_depth_mm, 
           color = species)) +
  geom_point()
```
:::

::: {.column width="60%"}

Something else

:::

::::

::: {.notes}
I may write a lot of extra stuff here in case you are interested.
:::


## Title


```{r}
#| echo: true

ggplot(penguins,                        
       aes(x = bill_length_mm,          
           y = bill_depth_mm, 
           color = species)) +
  geom_point()
```


## This prints to Comparison

:::: {.columns}

::: {.column}

```{r}
#| echo: true
#| output: true

plot(penguins$bill_depth_mm, 
     penguins$bill_length_mm)

```

:::

::: {.column}

```{r}
#| echo: true
#| output: true

ggplot(penguins,                        
       aes(bill_depth_mm,          
           bill_length_mm)) +
  geom_point()
```

:::

::::

--------------------------------------------------------------------------------

## Using `with()` in *base R*

:::: {.columns}

::: {.column}

```{r}
#| echo: true
#| output: true

with(penguins,
     plot(bill_depth_mm, 
          bill_length_mm))

```

:::

::: {.column}

```{r}
#| echo: true
#| output: true

ggplot(penguins,                        
       aes(bill_depth_mm,          
           bill_length_mm)) +
  geom_point()
```

:::

::::


--------------------------------------------------------------------------------

## Changing the `ggplot` theme

:::: {.columns}

::: {.column}

```{r}
#| echo: true
#| output: true

with(penguins,
     plot(bill_depth_mm, 
          bill_length_mm))

```

:::

::: {.column}

```{r}
#| echo: true
#| output: true

ggplot(penguins,                        
       aes(bill_depth_mm,          
           bill_length_mm)) +
  geom_point() +
  theme_classic()
```

:::

::::


--------------------------------------------------------------------------------

## *Some* further styling options

:::: {.columns}

::: {.column}

```{r}
#| echo: true
#| output: true

par(mar=c(4.5, 4.2, 1, 7), xpd=TRUE)          # set custom plot size

with(penguins,                                # select data frame
     plot(bill_depth_mm,                      # select variables to plot
          bill_length_mm,
          col = species,                      # grouping colour
          bty = "L")                          # change plot border style 
     )                                                            

with(penguins,
      legend("right",                         # legend position
             inset=c(-0.23,0.5),              # legend custom location (outside the plot)
             legend = levels(species),        # legend labels
             pch = 16,                        # legend symbol size
             col = unique(species),           # legend symbol colour
             bty = "n"                        # remove legend borders
             )
     )

```


:::

::: {.column}

```{r}
#| echo: true
#| output: true

ggplot(penguins,                              # draw empty canvas
       aes(bill_depth_mm,                     # select variables to plot
           bill_length_mm,
           color = species)) +                # grouping colour
  geom_point() +                              # select plot type
  theme_classic()                             # select pre-set style settings (theme)
```

:::

::::






--------------------------------------------------------------------------------


## Test


```{r}
#| echo: true
#| output-location: column-fragment
#| code-line-numbers: "|2"

library(ggplot2)

mtcars |> 
  ggplot(aes(x = disp, y = mpg)) +
  geom_point() +
  geom_smooth(method = "loess", formula = "y~x")
```




--------------------------------------------------------------------------------


## Final


This is an [important sentence!]{.fragment .highlight-red}



