---
title: "Interaction models"
subtitle: "Week 4"
format:
  pptx:
    incremental: true   
    reference-doc: templates/SOC2069_pptx_template.pptx
---

## Statistical models

-   **(general) linear models**

    -   For when we are interested in the relationship of some ***continuous*** outcome variable, say $Y$, and one or more explanatory (predictor) variables, say $X_1 \dots {p}$
        -   for $p=1$. we have a *simple* (bi-variate) regression;
            -   a more general version of a ***correlation*** between two variables, using the true, unstandardised scales on which variables are measured;
            -   when driven by substantive theory on a specific topic we wish to explain the observed variation in one of the variables (designated as the "dependent"/"outcome"/"explained" variable) based on the variability of the other variable (the "independent"/"predictor"/"explanatory" variable)
        -   for $p>1$, we have a *multiple* regression;
            -   a more general version of a ***partial correlation*** between two variables in the presence of other covariates;
    -   Lets us characterize the relationship between $Y$ and $X$ as: $Y = \beta_0 + \beta_1 X + \epsilon$, where
        -   $\beta_0$ is just a baseline, average value of $Y$ when we assign $X$ the value **0** (so exclude its effect)
        -   $\beta_1$ is the predicted difference in the value of $Y$ associated with a *one unit* change in $X$; where "unit" depends on how $X$ is measured (e.g. one year of age; one extra £ of income; a 1-point jump on an attitudinal scale running from 0 to 10; being 1 (=Female) compared to 0(=Male);
        -   the $\epsilon$ is "noise"; accounts for the deviations between the data points and the expected linear trend line
        -   $\epsilon$ must follow a ***normal distribution*** to satisfy the mathematical requirements for a linear model; if not, we need to think about other modelling approaches

-   **Generalised linear models**

    -   Logistic regression
        -   For when we are interested in the relationship between a **binary** outcome variable and one or more explanatory variables

## Interaction models

-   In linear models, each explanatory variable is assumed to have an effect on the outcome variable that is independent of the values of other explanatory variables included in the model  main effects
-   We can reject this assumption and model any theoretically motivated dependencies between the explanatory variables themselves
-   We "multiplicate" the effect of two explanatory variables to estimate their effect as dependent on the values of the other explanatory variable(s)

## Remember...

```{r, include=FALSE}
# Packages and data

pacman::p_load(tidyverse,             # metapackage; general data management (especially the `dplyr` package)
               easystats,             # metapackage; data management and model summaries
               gtsummary,             # data and model summary tables
               ggformula,             # formula syntax to simplify `ggplot2` graphics
               sjlabelled,            # work with labelled survey data from other statistical packages
               ggeffects,             # extract and plot marginal/conditional/interaction effects from models
               marginaleffects,       # extract and plot marginal/conditional/interaction effects from models
               latex2exp,             # math annotation on graphs
               betareg,               # Beta regression for proportional data
               ggpubr,                # GGplot functions (combine and arrange plots on page)
               patchwork              # same
               )                


inequality <- data_read("https://github.com/CGMoreh/SOC2069/raw/main/Data/for_analysis/lab3macro.rds")
```

```{r}
# Data wrangling and extracting model components

# remove missing, arrange by Region, set up graphing aesthetics

selected_countries <- c("United Kingdom", "Nicaragua") 

inequality <- inequality |> 
  drop_na(trust_pct, s80s20) |> 
  arrange(Region) |> 
  mutate(
    trust_prop = structure(trust_pct/100, label = "Prop. people agreein that 'most people can be trusted'"),
    non_trust_prop = structure(1-trust_prop, label = "Prop. people agreeing that 'one cannot be too careful'"),
    # aesthetics
    colours = case_when(country == selected_countries[1] ~ "#009C6B",
                        country == selected_countries[2] ~ "#FF8916",
                        TRUE ~ "black"),
    sizes = case_when(country %in% selected_countries ~ 2.5, 
                      TRUE ~ 1.5),
    transparency = case_when(country %in% selected_countries ~ 1,
                             TRUE ~ 0.1)
  )
  

mod <- lm(trust_pct ~ s80s20, data = inequality)


inequality <- inequality |> 
  data_modify(fitted = case_when(country %in% selected_countries ~ mod$fitted.values))



ggeffects::ggpredict(mod, terms = c("s80s20")) |> plot()

model_parameters(mod)

intercept <-  coef(mod)[1] |> as.numeric() |> round(1)

pred <- estimate_prediction(mod) |> 
  mutate(country = inequality$country,
         coef = mod$effects) |> 
  relocate(country) |> 
  filter(country %in% selected_countries)






```

```{r}
# Simple scatter with fit line

gf_point(trust_pct ~ s80s20, data = inequality) |> 
  gf_lm(color="red", size = 1) |> 
  gf_text(label = ~ country, size = 2.5, hjust = 0, nudge_x = 0.15, nudge_y = 0.15) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 80), breaks = c(seq(0, 80, 10))) +
  scale_x_continuous(limits = c(0, 17)) 
  
```

```{r echo=FALSE, warning=FALSE}



gf_point(trust_pct ~ s80s20, alpha = inequality$transparency, color = inequality$colours, size = inequality$sizes, data = inequality) |> 
  gf_lm(color="darkred", linewidth = 1) |> 
  gf_text(label = ~ country, size = 3, hjust = 0, nudge_x = 0.15, nudge_y = 0.15, alpha = inequality$transparency, color = inequality$colours) |> 
  # gf_hline(yintercept = fit, color = "darkblue", linetype = "dashed") |> 
  gf_hline(yintercept = 0, color = "black") |> 
  gf_vline(xintercept = 0, color = "black") |> 
  gf_segment(fitted + trust_pct ~ s80s20 + s80s20, linetype = "solid", color = inequality$colours, size = 0.7) |> 
  gf_segment(fitted + -10 ~ s80s20 + s80s20, linetype = "dashed", color = inequality$colours, size = 0.6) |>
  gf_segment(fitted + fitted ~ -1 + s80s20, linetype = "dashed", color = inequality$colours, size = 0.6) +
  theme_bw() +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_y_continuous(limits = c(-10, 80), breaks = c(seq(-10, 80, 10)), expand = c(0,0)) +
  scale_x_continuous(limits = c(-1, 17), breaks = c(seq(-1, 17, 1)), expand = c(0,0)) +
  annotate("point", shape = 4, x = 0, y = intercept, size = 3, color = "darkred", stroke = 2) +
  annotate("text", x = 1.7, y = intercept+3.4, size = 3, color = "black", label = paste(TeX(r"($\beta_0$)"), " (Intercept) ==", intercept), parse = TRUE) +
  annotate("text", x = 16.5, y = 75, size = 4, hjust = "right", label = paste("Model: ", TeX(r"($Trust = \beta_0 + \beta_1 \times Inequality + \epsilon$)")), parse = TRUE, color = "darkred") +
  annotate("text", x = 16.5, y = 67, size = 3, hjust = "right", label = paste(TeX(r"($Trust_{UK} \phantom{0} =  45.4 + (-3.1) \times \phantom{0} 5.40$ = 28.7)")), parse = TRUE, color = "#009C6B") +
  annotate("text", x = 16.5, y = 60, size = 3, hjust = "right", label = paste(TeX(r"($Trust_{Nic.} =  45.4 + (-3.1) \times 10.2$ = 13.7)")), parse = TRUE, color = "#FF8916")
  
```

```{r}

gf_point(trust_pct ~ s80s20, data = inequality, color = ~ Region) |> 
  gf_lims(y = c(1, 80)) |> 
  gf_lm() +
  theme_minimal()

```
