
# European Values Study, 2017 (EVS2017)

This dataset comes from the [European Values Study 2017 (EVS 2017)](https://search.gesis.org/research_data/ZA7500?doi=10.4232/1.13897). The raw data (Version 5.0.0, in SPSS format) and documentation was downloaded from the GESIS repository to a local folder, `Data/raw/evs2017/`:


```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false

fs::dir_info("../Data/raw/evs2017")[, c(1,3)] |> flextable::qflextable() |> flextable::fontsize(size = 10)

```

The SOC2069 version of the dataset can be reproduced from the raw data file with the `R` code shown in the expandable field below.

```{r}
#| warning: false
#| output: false
#| code-fold: true
#| code-summary: "Show the code"

## Packages (install first if needed) ----------------------------------------------------------------------------------------

library(here)           # Write paths to data relative to project directory on both Win and Mac
library(dplyr)          # Data wrangling
library(datawizard)     # All data import, export and wrangling
library(sjlabelled)     # Data import and label management

## Path to file --------------------------------------------------------------------------------------------------------------

datafile <- here("Data", "raw", "evs2017", "ZA7500_v5-0-0.sav.zip")

## Read in the data file -----------------------------------------------------------------------------------------------------

evs2017 <- sjlabelled::read_spss(datafile) |> 
  drop_labels() |>
  mutate(across(where(~ is.factor(.x) && nlevels(.x) > 9), 
                as_numeric),                                                    # Turn long factors to labelled numeric
         # across(c(), 
                # as_factor)                                                    # Long factors that should stay categorical
         ) |>                                      
  data_select(-c(studyno:fw_end, c_abrv, mode:fmissings, v228b:v229,
                 v231b, v231b_r, v233b, v233b_r, v242_r:v243_8cat,
                 v243_cs_DE1:v243_cs_GB2,v245:v248a,
                 v251b_r:v252_8cat, v252_cs_DE1:v252_cs_GB2,
                 v254:v258, v262_edulvlb:v262_8cat,
                 v262_cs_DE1:v263_8cat, v263_cs_DE1:v243_r_weight,
                 starts_with("f"),
                 ends_with("_IT", "_DK", "_cs", "_SE", "a", "b", "c"))
              )

## Save the dataset for SOC2069 --------------------------------------------------------------------------------------

saveRDS(evs2017, "for_analysis/evs2017.rds", compress = "bzip2")

```

If you use the EVS2017 data in your work, cite it as follows:

 > EVS (2022). European Values Study 2017: Integrated Dataset (EVS 2017). GESIS, Cologne. ZA7500 Data file Version 5.0.0, https://doi.org/10.4232/1.13897.
 
 
 
## EVS2017 variable search

```{r}
#| echo: false
#| message: false
#| warning: false
#| eval: true


d <- readRDS("for_analysis/evs2017.rds") |> remove_empty_columns()

Levels <- c()
for (i in names(d)) {
  Levels <- c(Levels, nlevels(as_factor(d[[i]])))
}
  
searchtable <- data_codebook(d) |>  
  select(-c(1, 6:10)) |> 
  filter(Name != "") |> 
  rename(`N (%) missing` = Missings) |>
  mutate(Levels = Levels) |> 
  relocate(Levels, .after = Type)

searchtable |> datatable(rownames = TRUE,
            extensions = 'FixedHeader',
            options = list(columnDefs = list(list(width = "50%", targets = "Label")),
                           pageLength = 10,
                           lengthMenu = c(5, 20, 50, nrow(searchtable)),
                           fixedHeader = TRUE,
                           headerCallback = DT::JS(                               # https://github.com/rstudio/DT/issues/783
                              "function(thead) {",
                              "  $(thead).css('font-size', '0.9em');",
                              "}"
                            ))) |>
  formatStyle(columns = colnames(searchtable), fontSize = '85%')

```

 
 
 
 

## EVS2017 codebook

<details>
  <summary><i>View the full codebook</i></summary>

```{r}
#| echo: false
#| message: false
#| warning: false

`SOC2069: European Values Study, 2017` <- d

`SOC2069: European Values Study, 2017` |>  
    data_codebook(variable_label_width = 50, 
                  value_label_width = 45,
                  max_values = 10,
                  range_at = 5) |> 
    print_html(font_size = "80%",
               line_padding = 0)
```

</details>
