
# European Values Study (2017)

This dataset comes from the [European Values Study 2017 (EVS 2017)](https://search.gesis.org/research_data/ZA7500?doi=10.4232/1.13897). The data (Version 5.0.0) was downloaded in SPSS format from the GESIS repository and save to the `raw/evs2017/` folder:


```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false

fs::dir_info("raw/evs2017")[, c(1,3)] |> gt::gt_preview() |> gt::cols_align(align = "left") |> gt::tab_options(table.width = "70%")

```

The SOC2069 version of the dataset can be reproduced from the raw data file with the `R` code shown in the expandable field below.

```{r}
#| warning: false
#| output: false
#| code-fold: true
#| code-summary: "Show the code"

## Packages (install first if needed) ----------------------------------------------------------------------------------------

library(here)           # Write paths to data relative to project directory on both Win and Mac
library(dplyr)          # Data wrangling
library(datawizard)     # All data import, export and wrangling
library(sjlabelled)     # Data import and label management

## Path to file --------------------------------------------------------------------------------------------------------------

datafile <- here("Data", "raw", "evs2017", "ZA7500_v5-0-0.sav.zip")

## Read in the data file -----------------------------------------------------------------------------------------------------

evs2017 <- sjlabelled::read_spss(datafile) |> 
  drop_labels() |>
  mutate(across(where(~ is.factor(.x) && nlevels(.x) > 9), 
                as_numeric),                                                    # Turn long factors to labelled numeric
         # across(c(), 
                # as_factor)                                                    # Long factors that should stay categorical
         ) |>                                      
  data_select(-c(studyno:fw_end, c_abrv, mode:fmissings, v228b:v229,
                 v231b, v231b_r, v233b, v233b_r, v242_r:v243_8cat,
                 v243_cs_DE1:v243_cs_GB2,v245:v248a,
                 v251b_r:v252_8cat, v252_cs_DE1:v252_cs_GB2,
                 v254:v258, v262_edulvlb:v262_8cat,
                 v262_cs_DE1:v263_8cat, v263_cs_DE1:v243_r_weight,
                 starts_with("f"),
                 ends_with("_IT", "_DK", "_cs", "_SE", "a", "b", "c"))
              )

## Save the dataset for SOC2069 --------------------------------------------------------------------------------------

saveRDS(evs2017, "for_analysis/evs2017.rds", compress = "bzip2")

```

If you use the EVS2017 data in your work, cite it as follows:

 > EVS (2022). European Values Study 2017: Integrated Dataset (EVS 2017). GESIS, Cologne. ZA7500 Data file Version 5.0.0, https://doi.org/10.4232/1.13897.

## EVS2017 Codebook

```{r}
#| echo: false
#| message: false
#| warning: false


`SOC2069: European Values Study, 2017` <- evs2017

`SOC2069: European Values Study, 2017` |> 
    data_codebook(variable_label_width = 50, 
                  value_label_width = 25,
                  max_values = 10,
                  range_at = 5) |> 
    print_html(font_size = "80%",
               line_padding = 0)
```