
# International Social Survey Programme, 2020 (ISSP2020)

This dataset comes from the [International Social Survey Programme (2020)](https://www.gesis.org/en/issp/modules/issp-modules-by-topic/environment/2020). The data (version 2.0.0) was downloaded in SPSS format to a local folder, `raw/issp2020/`:

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false

# The folder, which also contains the original data documentation, looks like this:

fs::dir_info("raw/issp2020")[, c(1,3)] |> gt::gt_preview() |> gt::cols_align(align = "left") |>
  gt::tab_options(table.width = "70%")

```

The SOC2069 version of the dataset can be reproduced from the raw data file with the `R` code shown in the expandable field below.

```{r}
#| warning: false
#| output: false
#| code-fold: true
#| code-summary: "Show the code"

## Packages (install first if needed) ----------------------------------------------------------------------------------------

library(here)           # Write paths to data relative to project directory on both Win and Mac
library(dplyr)          # Data wrangling
library(datawizard)     # All data import, export and wrangling
library(sjlabelled)     # Data import and label management

## Path to file --------------------------------------------------------------------------------------------------------------

datafile <- here("Data", "raw", "issp2020", "ZA7650_v2-0-0.sav")

## Read in the data file -----------------------------------------------------------------------------------------------------

issp2020 <- read_spss(datafile) |> 
  drop_labels() |> 
  mutate(across(where(~ is.factor(.x) && nlevels(.x) > 9), 
                as_numeric),                                                    # Turn long factors to labelled numeric
         across(c(country, c_alphan, v16, ISCO08, RELIGGRP),
                as_factor)) |>                                                  # Long factors that should stay categorical
  set_labels(BIRTH, AGE, WRKHRS, NSUP, HOMPOP, HHCHILDR,  HHTODD,
             labels = "") |>                                                    # Pure numeric, no labels needed
  data_select(-c(studyno:doi, c_sample, CASEID:PARTIALS,                        # Technical variables
                 AT_ISCD:ISCED, AT_RELIG:ZA_RELIG, AT_PRTY:ZA_PRTY,
                 AT_ETHN1:ZA_ETHN1, AT_RINC:ZA_INC, AT_REG:ZA_REG,              # Country specific variables
                 SPWORK:SPMAINST, TOPBOT, HHADULT, F_BORN, M_BORN))             # Too much detail


## Save the dataset for SOC2069 --------------------------------------------------------------------------------------

# rds
saveRDS(issp2020, "for_analysis/issp2020.rds", compress = "bzip2")

```


If you use the ISSP2020 data in your work, cite it as follows:

  > ISSP Research Group (2023). International Social Survey Programme: Environment IV - ISSP 2020. GESIS, Cologne. ZA7650 Data file Version 2.0.0, https://doi.org/10.4232/1.14153.



## ISSP2020 Codebook

```{r}
#| echo: false
#| message: false
#| warning: false


`SOC2069: International Social Survey Programme, 2020` <- issp2020 

`SOC2069: International Social Survey Programme, 2020`  |> 
    data_codebook(variable_label_width = 50, 
                  value_label_width = 45,
                  max_values = 11,
                  range_at = 6) |> 
    print_html(font_size = "80%",
               line_padding = 0)
```




