
# International Social Survey Programme, 2020 (ISSP2020)

This dataset comes from the [International Social Survey Programme (2020)](https://www.gesis.org/en/issp/modules/issp-modules-by-topic/environment/2020). The raw data (version 2.0.0, in SPSS format) and documentation was downloaded to a local folder, `Data/raw/issp2020/`:

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false

# The folder, which also contains the original data documentation, looks like this:

fs::dir_info("../Data/raw/issp2020")[, c(1,3)] |> flextable::qflextable() |> flextable::fontsize(size = 10)

```

The SOC2069 version of the dataset can be reproduced from the raw data file with the `R` code shown in the expandable field below.

```{r}
#| warning: false
#| output: false
#| code-fold: true
#| code-summary: "Show the code"

## Packages (install first if needed) ----------------------------------------------------------------------------------------

library(here)           # Write paths to data relative to project directory on both Win and Mac
library(dplyr)          # Data wrangling
library(datawizard)     # All data import, export and wrangling
library(sjlabelled)     # Data import and label management

## Path to file --------------------------------------------------------------------------------------------------------------

datafile <- here("Data", "raw", "issp2020", "ZA7650_v2-0-0.sav")

## Read in the data file -----------------------------------------------------------------------------------------------------

issp2020 <- read_spss(datafile) |> 
  drop_labels() |> 
  mutate(across(where(~ is.factor(.x) && nlevels(.x) > 9), 
                as_numeric),                                                    # Turn long factors to labelled numeric
         across(c(country, c_alphan, v16, ISCO08, RELIGGRP),
                as_factor)) |>                                                  # Long factors that should stay categorical
  set_labels(BIRTH, AGE, WRKHRS, NSUP, HOMPOP, HHCHILDR,  HHTODD,
             labels = "") |>                                                    # Pure numeric, no labels needed
  data_select(-c(studyno:doi, c_sample, CASEID:PARTIALS,                        # Technical variables
                 AT_ISCD:ISCED, AT_RELIG:ZA_RELIG, AT_PRTY:ZA_PRTY,
                 AT_ETHN1:ZA_ETHN1, AT_RINC:ZA_INC, AT_REG:ZA_REG,              # Country specific variables
                 SPWORK:SPMAINST, TOPBOT, HHADULT, F_BORN, M_BORN))             # Too much detail


## Save the dataset for SOC2069 --------------------------------------------------------------------------------------

# rds
saveRDS(issp2020, "for_analysis/issp2020.rds", compress = "bzip2")

```


If you use the ISSP2020 data in your work, cite it as follows:

  > ISSP Research Group (2023). International Social Survey Programme: Environment IV - ISSP 2020. GESIS, Cologne. ZA7650 Data file Version 2.0.0, https://doi.org/10.4232/1.14153.



## ESS10 variable search

```{r}
#| echo: false
#| message: false
#| warning: false
#| eval: true


d <- readRDS("for_analysis/issp2020.rds") |> select(-ZA_ETHN2)                        # empty!

Levels <- c()
for (i in names(d)) {
  Levels <- c(Levels, nlevels(as_factor(d[[i]])))
}
  
searchtable <- data_codebook(d) |>  
  select(-c(1, 6:10)) |> 
  filter(Name != "") |>
  rename(`N (%) missing` = Missings) |>
  mutate(Levels = Levels) |>
  relocate(Levels, .after = Type)

searchtable |> datatable(rownames = TRUE,
            extensions = 'FixedHeader',
            options = list(columnDefs = list(list(width = "50%", targets = "Label")),
                           pageLength = 10,
                           lengthMenu = c(5, 20, 50, nrow(searchtable)),
                           fixedHeader = TRUE,
                           headerCallback = DT::JS(                               # https://github.com/rstudio/DT/issues/783
                              "function(thead) {",
                              "  $(thead).css('font-size', '0.9em');",
                              "}"
                            ))) |>
  formatStyle(columns = colnames(searchtable), fontSize = '85%')

```







## ISSP2020 codebook

<details>
  <summary><i>View the full codebook</i></summary>

```{r}
#| echo: false
#| message: false

`SOC2069: International Social Survey Programme, 2020` <- d

`SOC2069: International Social Survey Programme, 2020` |>  
    data_codebook(variable_label_width = 50, 
                  value_label_width = 45,
                  max_values = 11,
                  range_at = 6) |> 
    print_html(font_size = "80%",
               line_padding = 0)
```

</details>


