---
title: "Worksheet 10"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    toc_highlight: yes
    tabset: yes
    number_sections: no
    anchor_sections: no
    self_contained: yes
    code_download: no
    highlight: default
    css: ./../my_themes/labcss.css
---

```{r, include=FALSE, purl=FALSE}
options(repos = list(CRAN="http://cran.rstudio.com/"))
```

```{r setup, include = FALSE, purl=FALSE}
# knitr::knit_hooks$set(purl = knitr::hook_purl)

knitr::opts_chunk$set(error = FALSE, warning=FALSE, comment="", prompt=FALSE, message = F, strip.white = F, fig.width = 5, fig.asp = .8, fig.align = "center", fig.show = 'hold', class.output = "outputbox")
```

```{r klippy, echo=FALSE, include=TRUE, purl=FALSE}
pacman::p_load_gh("rlesur/klippy"); klippy(position = c('right'))
```

------------------------------------------------------------------------

    Timetable week: 15
    Topic: "Presenting quantitative results"

------------------------------------------------------------------------

# Intro

The aim of this final week is to gain some practice in getting the various outputs we have produced as part of a statistical analysis onto "paper", as part of the process of writing up our findings into a report.

In every quantitative analysis, we need to ensure that our analysis is open and reproducible. This is increasingly a professional standard expected of all data analysts in the social sciences. This means that we need to have an efficient way in which to share our **analysis code** and well as our **outputs** and our **interpretations** of our findings with our readers. This is also what the **assignment** requires.

`RStudio` has an efficient way of handling this requirement with the use of the **R markdown** document format, ***.Rmd*** (a more recent version of which is the *Quarto* document type, ***.qmd***, but we'll stick with `.Rmd` for now).

In fact, these worksheets themselves were written in `R Markdown`, so you have already witnessed how the output may look like: we have this text you are reading right now, and we can also include and execute `R` code chunks such as this one:

````{verbatim, purl=FALSE}
```{r}
# Let's first (install and) load the packages we want to use.

install.packages("tidyverse")
install.packages("mosaic")
install.packages("sjmisc")
install.packages("jtools")

library(tidyverse)
library(mosaic)
library(sjmisc)
library(jtools)
```
````


While one can do amazing things with `.Rmd` documents and there are various settings that one can use to customise the output in every way possible and in various output formats, we'll only cover the very basics that allow you to complete a full data analysis project (**such as your Assignment 2!!**) in an `.Rmd` document.

# Readings

Instead of readings, you have some online training to complete this week:

-   Component 6 of "R for Social Scientists": <https://datacarpentry.org/r-socialsci> ("Getting started with R Markdown", around 1 hour).

And/or the following:

-   Official "R Markdown" tutorial from RStudio: <https://rmarkdown.rstudio.com/lesson-1.html>
-   Andy Field's "Getting started in R and RStudio" (see on the software tutorial page too), especially the section on 'R Markdown' (<http://milton-the-cat.rocks/learnr/r/r_getting_started/#section-r-markdown>)

Further training:

For those interested in developing their `R` skills in creating publishable-quality visualisations (graphs, figures)

-   Component 5 of "R for Social Scientists": <https://datacarpentry.org/r-socialsci> ("Data Visualisation with ggplot2", around 2 hours).

Secondary readings:

-   Neuman (2014): Chapter 15 (pp. 513--546)


# Exercise 1: Open and get to know an `.Rmd` document

`About 30  minutes`

------------------------------------------------------------------------

1.  As usual, **open the `R Studio` interface** by clicking on the *SOC2069-Statistical-analysis**.Rproj*** file included in the ***SOC2069-Statistical-analysis*** project folder that you downloaded from Canvas in Lab6. The folder should be stored on your Newcastle University **OneDrive** and accessible from any computer

    `r fontawesome::fa("wrench")` If you haven't yet downloaded the project folder in TW11 (Lab6), then [download it from Canvas](https://ncl.instructure.com/courses/48074/pages/data-access-and-description?module_item_id=2544656).

2.  ***Not*** as usual, instead of opening a new `R` script for the session, open a new "R Markdown" document and give it the Title "Rmd practice" in the pop-up window (you can also add your name if you want, and the current date is included automatically, but you can change or delete that if you want):


<center>

<video width="320" height="240" controls><source src="images/Open_rmd_1.mp4" type="video/mp4"></video>

or

<video width="320" height="240" controls><source src="images/open_rmd_2.mp4" type="video/mp4"></video>

</center>


When you open a new `R Markdown` document from the Menu, the document that opens up is actually a micro-tutorial showing you the basic elements of the document format. Read through the included information and then "knit" the document in both *html* and *Microsoft Word* format to see the results:


<center>

To get a Microsoft Word document as output:

<video width="320" height="240" controls><source src="images/knit_rmd_word.mp4" type="video/mp4"></video>

To get an HTML document as the output:

<video width="320" height="240" controls><source src="images/knit_rmd_html.mp4" type="video/mp4"></video>

</center>

The reason why you are getting outputs (the table and the figure) is that the document uses two dataset included by default in `R` (one called "cars", the other "pressure"), so you don't need to first load the data. Of course, for your own work, you will need to include the same code we have usually used in order to load the UKHLS dataset and complete your analysis.

You can compare the contents of the `.Rmd` document to those in the "knitted" output. Notice that lines starting with one or more hashtags (#) are transformed into section headings. For your assignment, you can use one hashtag (#) for the main title of your report, then two hashtags (##) for subsections like Introduction, Methods, etc. Further hashtags mean further heading levels, but for a very short essay/report, you don't need more than two levels.

::: {.taskbox .task}

**Edit the *R Markdown* practice document**

For this quick exercise, let's use the same "cars" dataset that is included with `R` and whose summary is already on the sheet.The dataset contains two variables that give the ***speed*** of cars and the ***dist***ances taken to stop, recorded in the 1920s. (Type `?cars`  in the Console to open up a Help window briefly explaining where these data are from).

- select and delete all the text under the heading "## Including Plots"
- create a new code chunk under "## Including Plots"
- inside the code chunk, write the command for a simple linear regression that regresses "dist" (the outcome) on "speed" (the predictor)
- **you can run the commands in a code chunk to see the results by clicking on the *green arrow* in the upper-right corner of the chunk**
- In a new line under the output, write a brief sentence about the result: what is the relationship between a car's speed (measured in mph) and the distance it requires to stop (measured in ft)
- finally, "knit" the document to Microsoft Word to see the results (**Important**: first, close the Word document from the previous knit if it's still open, otherwise it cannot be overwritten and the knitting process will fail)

```{r, eval=FALSE, echo=FALSE}

# This is enough to complete the student task

lm(dist ~ speed, data = cars)
```

:::


# Exercise 2: Complete your assignment today!

To make your life easier, you can download another "R Project" folder for the assignment from here:

The folder contains:
- the UKHLS dataset
- a template `.Rmd` document to use for the assignment, with some useful default settings included ("my_assignment_RENAME.Rmd")
- an example Microsoft Word document, the "knitted" output from the `.Rmd`

Open the "my_assignment_RENAME.Rmd" document, read through it, and begin working on your assignment, following the same steps we practiced in previous weeks.

The aim is to complete the data analysis part of the assignment today, in class. You can leave the literature review and discussion for later. But if you manage finish this exercise and export your results to Microsoft Word or to HTML, you won't even need to use `R Studio` outside class to complete the assignment. You can just add the additional content directly onto the Word document as you would do with any other essay. 

If you output to HTML instead, it should be easy to copy the output to a Word document or another word editor, for example:

<center>
<video width="320" height="240" controls><source src="images/copy-html.mp4" type="video/mp4"></video>
</center>

Ideally, of course, you would complete the entire analysis and writing on the `.Rmd` document in `R Studio`, including the literature, reference list, etc. and only "knit" the final version that you are ready to submit.

```{r eval=FALSE, include=FALSE, purl=FALSE}
# Just run this chunk manually at the end...
knitr::purl("Lab10.Rmd", documentation = 0)
```
